Almost compatible functions and infinite length games

### 2015-12-02 articles/almost-compatible-functions-and-topological-games/article.tex
cite weiss et al
change S to alcompS since S(kappa) means other things in other papers. more connections to splendid paper

  \begin{definition}
    The statement \(S(\kappa)\) (given as \(S(\kappa,\aleph_0,\omega)\)
    in \cite{MR1129143}) claims that there exist one-to-one functions
    \(f_A:A\to\omega\) for each \(A\in[\kappa]^{\leq\aleph_0}\) such that
    the collection \(\{f_A:A\in[\kappa]^{\leq\aleph_0}\}\) is pairwise
    almost compatible.
  \end{definition}

  In the same paper, Scheepers noted that \(S(\omega_1)\) holds in \(ZFC\),
  and that it's possible to force \(\mf c\) to be arbitrarily large
  while preserving \(S(\mf c)\). However, \(S(\mf c^+)\) always fails.
  This axiom may be applied to obtain a winning \(2\)-tactic for \(\pl F\)
  in the countable-finite game.

@@

  The relationship between \(\schFillStrictGame\kappa\) and
  \(\menGame{\oneptlind\kappa}\) is strong; in both games \(\pl C\) essentially
  chooses a countable subset of \(\kappa\) followed by \(\pl F\) choosing
  a finite subset of that choice, and it's easy to see the winning perfect
  information strategy for \(\pl F\) in both games.
  In addition, it was shown in
  \cite{clontzMengerGamePreprint} that when \(S(\kappa)\) holds,
  \(\pl F\) has a winning \(2\)-Markov strategy in
  \(\menGame{\oneptlind\kappa}\).

@@

  One might hope that \(\oneptlind{(\mf c^+)}\) might answer this question
  in the affirmative as \(S(\mf c^+)\) fails, but we will show that assuming
  \(V=L\), \(\pl F\kmarkwin{2}\menGame{\oneptlind\kappa}\) for all
  cardinals \(\kappa\).

@@

  We may weaken Scheeper's \(S(\kappa)\) as follows:

  \begin{definition}
    The statement \(S'(\kappa)\) weakens \(S(\kappa)\) be only requiring
    the witness functions \(f_A:A\to\omega\) to be finite-to-one.
  \end{definition}

  The following observation will be convenient.

  \begin{proposition}
    \(S(\kappa)\) and \(S'(\kappa)\) need only be witnessed by
    functions \(\{f_A:A\in\mc S\}\) for some family \(\mc S\)
    cofinal in \([\kappa]^{\leq\aleph_0}\).
  \end{proposition}

@@

  In the next section we will show that \(S'(\kappa)\) is sufficient
  for the applications to the Scheepers and Menger games.
  In the meantime, we will demonstrate that \(S'(\kappa)\) is strictly
  weaker than \(S(\kappa)\).

@@

  \begin{theorem}
    \(S'(\kappa)\) holds whenever there exists a cofinal Kurepa family on \(\kappa\).
  \end{theorem}

@@

  We now construct a topology on \(\omega_n\) for each \(n<\omega\) which
  will witness a Kurepa family on \(\aleph_n\).

@@

@@

  \begin{proof}
    Note that the closure of any countable neighborhood is compact, and
    any Lindel\"of set is countable. This space is locally metrizable
    and thus first-countable, so cardinality and weight coincide for any
    subspace. The result is obvious if \(A\) is countable; otherwise
    let \(A=\{a_\alpha:\alpha<\omega_{n+1}\}\) and since basic neighborhoods
    are countable note any limit point is
    a limit point of \(A_\beta=\{a_\alpha:\alpha<\beta\}\)
    for some \(\beta<\omega_{n+1}\).
    Thus \(\cl A=\bigcup_{\beta<\omega_{n+1}}\cl{A_\beta}\) and
    by induction \(|\cl A|=|A|\).
  \end{proof}

@@

  \begin{theorem}
    For each $n\in \omega$, there is a \(T_2\), locally countable,
    $\omega$-bounded topology on $\omega_n$.
  \end{theorem}

@@

@@

  \begin{proof}
    We use the family \(\mc K\) of all compact
    open sets in the constructed topology on
    \(\omega_n\) as our witness. Of course, every Lindel\"of set in
    a locally countable space is countable, and the closure of every
    countable set is a compact countable set; thus \(\mc K\)
    is cofinal in \([\omega_n]^\omega\).
    It is Kurepa since for every countable set \(A\), there are only
    countably many distinct compact open sets in \(\cl A\): TODO finish
  \end{proof}

  This is alternatively a corollary of an observation of Todorcevic
  communicated by Dow in \cite{MR1229125}:
  if every Kurepa family of size at most \(\theta\)
  extends to a cofinal Kurepa family, then the same is true of \(\theta^+\).
  So the result follows as
  every Kurepa family \(\mc K\) of size \(\omega\) extends to
  the cofinal Kurepa family \([\bigcup\mc K]^\omega\).

  So we have our desired result.

  \begin{corollary}
    \(S'(\omega_n)\) holds for all \(n<\omega\). Under \(CH\), we have
    both \(S'(\omega_2)\) and \(\neg S(\omega_2)\).
  \end{corollary}

@@

  \begin{corollary}[\(V=L\)]
    \(S'(\theta)\) holds for all cardinals.
  \end{corollary}

  In particular, \(S(\omega_2)\) fails under \(CH\), showing the two are
  distinct. Actually, \(CH\) is not required to have \(S(\omega_2)\) fail.

@@

  On the other hand, it's also consistent that \(S'(\theta)\) can fail.

  \begin{theorem}
    There's a model where \(S'(\omega_\omega)\)
    fails.
  \end{theorem}

@@

  \begin{question}
    Is \(S'(\theta)\) equivalent to having a Kurepa family on \(\theta\)?
  \end{question}

@@

    &
    &
    \pl F \ktactwin2\schFillStrictGame\kappa \\
  };
  \path[>=latex,->]
    (m-2-1) edge (m-1-2)
    (m-2-1) edge (m-2-2)
    (m-2-1) edge (m-3-3)
    (m-1-2) edge (m-2-2)
    (m-1-3) edge (m-2-3)
    (m-3-3) edge (m-2-3);
  \path[>=latex,<->]
    (m-1-2) edge (m-1-3)
    (m-2-2) edge (m-2-3);
\end{tikzpicture}
\end{center}
\caption{Diagram of Scheeper/Menger game implications with \(S'(\kappa)\)}
\label{GamesDiagram2}
\end{figure}

  \begin{theorem}
    Figure \ref{GamesDiagram2} holds. (Proven in [TODO cite])
    (Actually, TODO double-check that it works with just S', particularly
    the strict game)
  \end{theorem}

@@

  \begin{corollary}
    It is consistent that \(S'(\omega_\omega)\) fails,
    but as \(S'(\omega_k)\) holds for all \(k<\omega\), we have
    \(\pl F\kmarkwin2\schFillIntGame{\omega_\omega}\) and
    \(\pl F\kmarkwin2\schFillInitialGame{\omega_\omega}\).
  \end{corollary}

  A tricky topological question: does \(\pl F\win\menGame{X}\) imply
  \(\pl F\kmarkwin2\menGame{X}\)?
  (C showed that )
  Under \(V=L\), we cannot hope to find
  a counterexample using \(X=\oneptlind\kappa\) since
  \(S'(\kappa)\) and thus \(\pl F\kmarkwin2\schFillIntGame{\kappa}\)
  always holds.

>>>>

  \begin{definition}
    The statement \(\alcompS\kappa\) (given as \(S(\kappa,\aleph_0,\omega)\)
    in \cite{MR1129143} and \(S(\kappa)\) in \cite{clontzMengerGamePreprint})
    claims that there exist one-to-one functions
    \(f_A:A\to\omega\) for each \(A\in[\kappa]^{\leq\aleph_0}\) such that
    the collection \(\{f_A:A\in[\kappa]^{\leq\aleph_0}\}\) is pairwise
    almost compatible.
  \end{definition}

  In the same paper, Scheepers noted that \(\alcompS{\omega_1}\) holds in \(ZFC\),
  and that it's possible to force \(\mf c\) to be arbitrarily large
  while preserving \(\alcompS{\mf c}\). However, \(\alcompS{\mf c^+}\) always fails.
  This axiom may be applied to obtain a winning \(2\)-tactic for \(\pl F\)
  in the countable-finite game.

@@

  The relationship between \(\schFillStrictGame\kappa\) and
  \(\menGame{\oneptlind\kappa}\) is strong; in both games \(\pl C\) essentially
  chooses a countable subset of \(\kappa\) followed by \(\pl F\) choosing
  a finite subset of that choice, and it's easy to see the winning perfect
  information strategy for \(\pl F\) in both games.
  In addition, it was shown in
  \cite{clontzMengerGamePreprint} that when \(\alcompS{\kappa}\) holds,
  \(\pl F\) has a winning \(2\)-Markov strategy in
  \(\menGame{\oneptlind\kappa}\).

@@

  One might hope that \(\oneptlind{(\mf c^+)}\) might answer this question
  in the affirmative as \(\alcompS{\mf c^+}\) fails, but we will show that assuming
  \(V=L\), \(\pl F\kmarkwin{2}\menGame{\oneptlind\kappa}\) for all
  cardinals \(\kappa\).

@@

  We may weaken Scheeper's \(\alcompS{\kappa}\) as follows:

  \begin{definition}
    The statement \(\alcompSP{\kappa}\) weakens \(\alcompS{\kappa}\) be only requiring
    the witness functions \(f_A:A\to\omega\) to be finite-to-one.
  \end{definition}

  \begin{proposition}
    \(\alcompS{\kappa}\) and \(\alcompSP{\kappa}\) need only be witnessed by
    functions \(\{f_A:A\in\mc S\}\) for some family \(\mc S\)
    cofinal in \([\kappa]^{\leq\aleph_0}\).
  \end{proposition}

@@

  In the next section we will show that \(\alcompSP{\kappa}\) is sufficient
  for the applications to the Scheepers and Menger games.
  In the meantime, we will demonstrate that \(\alcompSP{\kappa}\) is strictly
  weaker than \(\alcompS{\kappa}\).

@@

  \begin{theorem}
    \(\alcompSP{\kappa}\) holds whenever there exists a cofinal Kurepa family on \(\kappa\).
  \end{theorem}

@@

  We now construct a topology on \(\omega_n\) for each \(n<\omega\) which
  will witness a Kurepa family of size \(\aleph_n\). A similar construction
  was previously shown in \cite{MR539228}.

@@

  We note that in
  \cite{MR539228} a \(T_2\) locally countable \(\omega\)-bounded space
  was called \term{splendid}, and the claim that there exists a splendid
  space of cardinality \(\kappa\) was called \(S(\kappa)\).

@@

  \begin{proof}
    Note that the closure of any countable neighborhood is compact, and
    any Lindel\"of set is countable. This space is locally metrizable
    and thus first-countable, so cardinality and weight coincide for any
    subspace. The result is obvious if \(A\) is countable; otherwise
    let \(A=\{a_\alpha:\alpha<\omega_{n+1}\}\) and since basic neighborhoods
    are countable note any limit point of \(A\) is
    a limit point of \(A_\beta=\{a_\alpha:\alpha<\beta\}\)
    for some \(\beta<\omega_{n+1}\).
    Thus \(\cl A=\bigcup_{\beta<\omega_{n+1}}\cl{A_\beta}\) and
    by induction \(|\cl A|=|A|\).
  \end{proof}

@@

  \begin{theorem}
    For each \(k< \omega\), there is a \(T_2\), locally countable,
    \(\omega\)-bounded topology on \(\omega_k\). (That is, \(S(\omega_k)\)
    for all \(k<\omega\)).
  \end{theorem}

@@

  \begin{lemma}
    The family of compact open sets in a locally countable, \(\omega\)-bounded
    topological space is a Kurepa family cofinal in \([X]^\omega\).
  \end{lemma}

  \begin{proof}
    Let \(\mc K\) collect all compact
    open subsets of \(X\). Of course, every Lindel\"of set in
    a locally countable space is countable, and the closure of every
    countable set is a compact countable set; thus \(\mc K\)
    is cofinal in \([X]^\omega\).
    It is Kurepa since every countable set is contained in a countable
    compact open subspace of \(X\); this subspace has a countable
    base of compact open sets, which closed under finite unions enumerates
    all compact open subsets of the subspace.
  \end{proof}

  The above lemma may alternatively be stated as ``\(S(\kappa)\) implies
      the existence of a cofinal Kurepa family on \(\kappa\)''.

@@

  This corollary is also an observation of Todorcevic
  communicated by Dow in \cite{MR1229125}:
  if every Kurepa family of size at most \(\theta\)
  extends to a cofinal Kurepa family, then the same is true of \(\theta^+\).
  So the result follows as
  every Kurepa family \(\mc K\) of size \(\omega\) extends to
  the cofinal Kurepa family \([\bigcup\mc K]^\omega\).

  With this we have our desired result.

  \begin{corollary}
    \(\alcompSP{\aleph_k}\) holds for all \(k<\omega\). Under \(CH\), we have
    both \(\alcompSP{\aleph_2}\) and \(\neg\alcompS{\aleph_2}\).
  \end{corollary}

@@

  \begin{corollary}[\(V=L\)]
    \(\alcompSP{\theta}\) holds for all cardinals.
  \end{corollary}

  In particular, \(\alcompS{\omega_2}\) fails under \(CH\), showing the two are
  distinct. Actually, \(CH\) is not required to have \(\alcompS{\omega_2}\) fail.

@@

  On the other hand, it's also consistent that \(\alcompSP{\theta}\) can fail.

  \begin{theorem}
    There's a model where \(\alcompSP{\omega_\omega}\)
    fails.
  \end{theorem}

@@

  \begin{question}
    Is \(\alcompSP{\theta}\) equivalent to having a Kurepa family on \(\theta\)?
  \end{question}

@@

    &
    &
    \pl F \ktactwin2\schFillStrictGame\kappa \\
  };
  \path[>=latex,->]
    (m-2-1) edge (m-1-2)
    (m-2-1) edge (m-2-2)
    (m-2-1) edge (m-3-3)
    (m-1-2) edge (m-2-2)
    (m-1-3) edge (m-2-3)
    (m-3-3) edge (m-2-3);
  \path[>=latex,<->]
    (m-1-2) edge (m-1-3)
    (m-2-2) edge (m-2-3);
\end{tikzpicture}
\end{center}
\caption{Diagram of Scheeper/Menger game implications with \(\alcompSP{\kappa}\)}
\label{GamesDiagram2}
\end{figure}

  \begin{theorem}[\cite{clontzMengerGamePreprint}]
    Figure \ref{GamesDiagram2} holds.
    (Actually, TODO double-check that it works with just S', particularly
    the strict game)
  \end{theorem}

@@

  \begin{corollary}
    It is consistent that \(\alcompSP{\omega_\omega}\) fails,
    but as \(\alcompSP{\omega_k}\) holds for all \(k<\omega\), we have
    \(\pl F\kmarkwin2\schFillIntGame{\omega_\omega}\) and
    \(\pl F\kmarkwin2\schFillInitialGame{\omega_\omega}\).
  \end{corollary}

  A tricky topological question: does \(\pl F\win\menGame{X}\) imply
  \(\pl F\kmarkwin2\menGame{X}\)?
  (C showed that )
  Under \(V=L\), we cannot hope to find
  a counterexample using \(X=\oneptlind\kappa\) since
  \(\alcompSP{\kappa}\) and thus \(\pl F\kmarkwin2\schFillIntGame{\kappa}\)
  always holds.
